---
title: "Population-Isoboles-DetailedSetup"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Population-Isoboles-DetailedSetup}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(populationIsoboles)
library(ggplot2)
if(!require("MASS")) {install.packages("MASS"); library(MASS)}
if(!require("dMod")) {install.packages("dMod"); library(dMod)}
.outputFolder <- "Output"
```

# Introduction

An R-script version of this vignette and the resources are available in the folder `system.file("examples/Example-1", package = "populationIsoboles")`

This vignette is ideal if you want to set up a complex simulation which requires an adaption of the `runPopulationIsoboles` function's internals.
It will hint to some extra details for the implementation

This vignette demonstrates how to use the fast isobole algorithm `fastIsoboles` in the context of population simulation of NLME models.
The underlying PKPD model describes malaria parasite clearance by two drugs acting jointly.
Implementing the model is very explicit in this vignette to showcase which functionality is needed for the population isobole simulation.
The relevant steps for model implementation are: 

1. Set up model for one parameter set
2. Set up simulation function with binary outputs: success = 1, failure = 0
3. Define function to simulate model for multiple parameter sets
4. Define objective function which evaluates success rates for multiple doses, given a parameter set

Implementing the parameter sampling has also been made explicit and presents the simplest form of Monte-Carlo sampling for NLME models.
The following five steps outline the parameter sampling procedure. 
Note that in this vignette, steps 2 and 3 were swapped, essentially meaning that in this case, the random effects have no uncertainty.

1. Define population parameters `theta` and the corresponding uncertainties encoded as covariance matrix `S`
2. Sample one population parameter vector based on the multivariate normal distribution parameterized by `theta` and `S`
3. Define covariance matrix of random effects `Omega`
4. Sample `Nsubj` realizations of random effects based on `Omega` and define "Individual" by adding the random effects to the fixed effects.
  In this case, random effects are added on log-scale
5. Repeat 2-4 for `Npop` populations.

The setup of the model is made this explicit, because the fast isoboles algorithm `fastIsoboles` needs to reuse the individual parameters.
Therefore, all stochastic sampling needs to be performed separately from the simulation.

The population isoboles simulation is shown explicitly to highlight the most relevant steps and arguments.
Finally, three different visualizations are shown.


# Load Model and Parameter information
```{r model}
# .. Model -----
model <- source(system.file(file.path("examples","Example-1","Resources","modelDeparsed.R"), package = "populationIsoboles"))$value
# .. Parameter information -----
parameterInfo <- source(system.file(file.path("examples","Example-1","Resources","parametersDeparsed.R"), package = "populationIsoboles"))$value
print(parameterInfo$theta[1:4]) # Theta = Point estimates of fixed effects
print(parameterInfo$S[1:4,1:4])     # S     = Covariance matrix for Theta
print(parameterInfo$Omega[1:4,1:4]) # Omega = Covariance matrix of random effects

# .. Show the basic PKPD model for one parameter -----
pars_doses <- c(AMTx1 = 500, AMTx2 = 500)
simtime    <- seq(0,24*28,24)
sim <- model(simtime, c(parameterInfo$theta, pars_doses))
plot(sim) +  
  geom_hline(yintercept = log(10), linetype = 2, color = "blue") + 
  annotate("text", 600, 3, label = "LLOQ", color = "blue") + 
  labs(x = "Time [h]", y  = "log Parasites / mL") + 
  guides(color = FALSE)
```

# Define functions to simulate populations
```{r populationSimulationFunctions}
# .. Parameter sampling -----

#' Sample realizations of population parameters
samplePopPars <- function(Nsubj, theta, S) {
  thetapars <- MASS::mvrnorm(1, mu = theta , Sigma = S)
  thetapars <- matrix(thetapars, nrow = Nsubj, ncol = length(theta), byrow = TRUE)
  thetapars
}

#' Sample realizations of random effects
sampleEtas <- function(Nsubj, Omega) {
  etas <- MASS::mvrnorm(Nsubj, mu = rep(0, dim(Omega)[1]) , Sigma = Omega)
  etas
}

#' Function to simulate individual parameters
#'
#' Draws a set of population parameters and adds random effects on log-scale
#'
#' @param Nsubj Numeric, number of subjects to sample
#' @param theta Population parameter vector
#' @param S     Covariance matrix for theta
#' @param Omega Covariance matrix for random effects
#'
#' @return matrix[individuals, parameters]
sampleIndividuals <- function(Nsubj, theta, S, Omega) {
  exp(log(samplePopPars(Nsubj,theta,S)) + sampleEtas(Nsubj, Omega))
}

#' Add dosing information to parameters
#'
#' Replicates parsIndiv for each dose combination
#'
#' @param AMTx1,AMTx2 Vectors of sample length with the respective doses
#' e.g. AMTx1[1] and AMTx2[1] are the first dose combination
#' @param parsIndiv output of [sampleIndividuals()]
#'
#' @return matrix with length(AMTx1) * nrow(parsIndiv) rows
expandParametersWithDosing <- function(AMTx1, AMTx2, parsIndiv) {
  doses_expanded <- matrix(rep(c(AMTx1,AMTx2), each = nrow(parsIndiv)), ncol = 2)
  doses_expanded <- `colnames<-`(doses_expanded, c("AMTx1", "AMTx2"))
  pars_expanded <- do.call(rbind, lapply(1:length(AMTx1), function(x) parsIndiv))
  pars <- cbind(doses_expanded, pars_expanded)
}


# .. Model simulation -----

#' simulate the model for one parameter set and decide if the patient is cured
#'
#' Cure criterion: log Parasite concentration BLOQ (< log10) at 28 days
#'
#' @param pars0
#'
#' @return logical(1L), TRUE = "cure", FALSE = "no cure"
simModelBinaryOutput <- function(pars0) {
  prediction <- (prd)(seq(0,24*28, 24), pars0, deriv = FALSE)[[1]]
  prediction <- prediction[nrow(prediction), "OUTPUT1", drop  = TRUE]
  cure28 <- prediction < log(10)
  cure28
}

#' Simulate model for a population
#'
#' @param pars matrix[individuals,c(dosingParameters, individualParameters)]
#'   Each row corresponds to one simulation, determined by all parameters in this row
#'   The columns of this matrix need to be all model parameters +  the dosing parameters AMTx1, AMTx2
#'
#' @return logical(nrow(individuals)) indicating if a subject is cured or not
simModelMultiplePars <- function(pars) {
  # Best to parallelize this step, but this created problems in the vignette building...
  iscured <- lapply(
    X = seq_len(nrow(pars)), FUN = function(i) {
      pars <- pars[i,,drop = TRUE]
      simModelBinaryOutput(pars)
    })
  do.call(c, iscured)
}

#' Final objective function to evaluate the cure rate
#'
#' @param AMTx1,AMTx2 Vectors of sample length with the respective doses
#'   e.g. AMTx1[1] and AMTx2[1] are the first dose combination
#' @param parsIndiv Output from [sampleIndividuals()]
#'
#' @return Vector of length(AMTx1), indicating the success rate at the dose
#'   combinations given by AMTx1,AMTx2
#' @export
objectiveFunction <- function(AMTx1,AMTx2, parsIndiv) {
  pars <- expandParametersWithDosing(AMTx1, AMTx2, parsIndiv)
  simres <- simModelMultiplePars(pars)
  cureRates <- vapply(seq_along(AMTx1), function(dose_i) {
    mean(simres[(dose_i-1)*nrow(parsIndiv) + 1:nrow(parsIndiv)])},
    FUN.VALUE = 0.1)
  cureRates
}

# .. Illustrate parameter sampling and model simulation -----
# Parameters of individual subjects of one population
set.seed(1234)
indivPars <- do.call(sampleIndividuals, c(list(N = 10), parameterInfo))
# Add dosing information, dose drug1 400mg, drug2 200 mg
pars <- cbind(AMTx1 = 500, AMTx2 = 500, indivPars)
print(simModelMultiplePars(pars)) # 4 out of 10 subjects are cured
# Evaluate objective function at two doses
AMTx1 <- c(500, 500)
AMTx2 <- c(300, 500)
print(objectiveFunction(AMTx1, AMTx2, indivPars))
```

# Run population isoboles
```{r}
# .. Additional Information for isobole simulation ----
# Determine maximum doses
# Need to have it in the format provided by the function below
gridInfo <- populationIsoboles:::gridInfo_default(AMTx1Max = 5000, AMTx2Max = 5000, imax = 5, offset = 0)
# Target cure rate in each population
objectiveValue <- 0.95 

# .. Population isobole simulation function -----
# Loop over populations:
# 1. Sample Individuals of this population
# 2. Run fastIsoboles, the fast isobole algorithm
# 3. Save results into .outputFolder/Simulations
run_PopulationIsoboles <- function(
  Npop, Nsubj,
  gridInfo, objectiveValue,
  objectiveFunction,
  sampleIndividuals,
  argsSampleIndividuals,
  .outputFolder
) {
  dir.create(file.path(.outputFolder, "Simulations"), recursive = TRUE)
  # Save gridInfo for later reuse. Need to save it under this exact name
  saveRDS(gridInfo, file.path(.outputFolder, "Simulations", "001-GridInfo.rds"))

  for (k in 1:Npop) {
    # Sample parameters
    parsIndiv <- do.call(sampleIndividuals, argsSampleIndividuals)
    # Define objective function which only takes arguments (AMTx1, AMTx2) and 
    #   has access to parsIndiv via the environment
    obj <- function(AMTx1, AMTx2) {
      objectiveFunction(AMTx1,AMTx2, parsIndiv)
    }
    # Run isobole algorithm for this population
    fastIsoboles(objfun = obj, objvalue = objectiveValue,
                    gridmin = gridInfo$gridmin, gridmax = gridInfo$gridmax,
                    imin = 5, imax = 5, FLAGverbose = FALSE, # Force 5 iterations
                    k = k, .outputFolder = .outputFolder)
    populationIsoboles:::PI_zip(populationIsoboles:::PI_files(.outputFolder, k)) # This step is useful if many populations are run. 
  }
}

# .. Run the population isoboles function -----
set.seed(1234)
Nsubj <- 20
run_PopulationIsoboles(
  Npop = 3, Nsubj = Nsubj, # Values should be higher to appropriately sample the parameter space.
                        # Needs powerful computer though
  gridInfo          = gridInfo,
  objectiveValue    = objectiveValue,
  objectiveFunction = objectiveFunction,
  sampleIndividuals = sampleIndividuals,
  argsSampleIndividuals = c(list(Nsubj = Nsubj), parameterInfo),
  .outputFolder     = .outputFolder
)
```

# Plot results

## Spaghetti plot of isoboles
```{r, Spaghetti,fig.width=7,fig.height=5}
plotSpaghetti(file.path(.outputFolder, "Simulations"))
```

## Spaghetti with confidence levels overlaid
```{r, Quantiles,fig.width=7,fig.height=5}
plotQuantiles(file.path(.outputFolder, "Simulations"))
```

## Confidence level response surface
```{r, CLRS,fig.width=7,fig.height=5}
plotCLRS(file.path(.outputFolder, "Simulations"))
```


# Clean up output folder
```{r}
unlink(.outputFolder, recursive = TRUE)
```



